---
title: "DS6306 - HW 4"
author: "Brandon Croom"
date: "May 30, 2019"
output: 
  html_document:
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#Load libraries to assist in HTML scraping and do cleanup. Pre-load the necessary librarys prior to question code for clarity.NOTE The bitops library is requred by RCurl.
library(XML)
library(bitops)
library(RCurl)
library(tidyr)
```

#Question 1
```{r}
#Load the IMBD Harry Potter webpage (http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1)
theUrl = "https://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1&_sm_nck=1"
URLResult = getURL(theUrl,.opts = list(ssl.verifypeer = FALSE) )

#read the HTML Tables from the webpage
AllHTMLTables = readHTMLTable(URLResult)

#Based on the HTML page design the Actor listing is the third table in the page. Extract out the list of actors and remove columns not needed. NOTE: This will also remove the "Rest of cast listed alphabetically" marker as it is in column 1
Actor_Listing = AllHTMLTables[[3]][,c(2,4)]

#Add variable names to the data frame
names(Actor_Listing) = c("Actor","Character")

#coalesce the Actor column to characters since it comes in as a factor
Actor_Listing$Actor = as.character(Actor_Listing$Actor)

#Remove observations with NA
Actor_Listing = na.omit(Actor_Listing)

#Replace the multiple spaces with a single space. NOTE: This will also collapse the Warwick Davis character names to "Griphook / Professor Filius Flitwick"
Actor_Listing$Character = gsub("\\s+"," ",Actor_Listing$Character)

#Split the Actor name into First Name and Surname columns. Find the last space in the Actor string value and split the Actor string
Actor_Listing = extract(Actor_Listing,Actor, into=c("FirstName","Surname"),"(.*)\\s+(\\S+)$")

#List out the first 10 rows
head(Actor_Listing,10)
```

***

#Question 2
```{r}
#Load the ESPN San Antonio Spurs webpage (http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs)
theURL = "http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs"
UrlResult = getURL(theURL,.opts = list(ssl.verifypeer = FALSE) )

#read the HTML Tables from the webpage
HTMLTable = readHTMLTable(UrlResult)

#Data analysis shows the sixth table is the player listing. Create a players data frame to store that data
PlayersDF = data.frame(HTMLTable[6])

#Split the players name and position into two distinct fields in the data frame
PlayersDF = extract(PlayersDF,Name, into=c("PlayerName","Position"),"(.*)\\s+(\\S+)$")

#Data analysis shows the eight table is the shooting statitics table. Create a shooting statistics data frame to store that data
ShootingStats = data.frame(HTMLTable[8])

#Combined the Field Goal percentage field from the Shooting Stats data frame into the players data frame
PlayersDF$FGPct = ShootingStats$ NULL.FG.

#Remove the last row of the data frame which is the totals row
PlayersDF = PlayersDF[-nrow(PlayersDF),]

#Field Goal percentage is a factor, convert to numeric for graphic purposes
PlayersDF$FGPct = as.numeric(paste(PlayersDF$FGPct))

#Set margins for the plot so things look nice
par(las=2)
par(mar=c(5,8,4,2))

#Build out the plot with the following parameters/conditions:
# - Set the x-axis limit to 100. NOTE: This helps prevent the legend from overlaying data until a player has 100% FG%
# - Make the barplot a horizontal plot for clarity
# - Since this is a horizontal bar plot, the Players Names will in effect be the labels for the y-axis.
# - Color code the bars based on player position. 
barplot(PlayersDF$FGPct, main="San Antonio Spurs Field Goal Percentage", ylab="",xlab="Field Goal Percentage",xlim=c(0,100),names.arg=c(PlayersDF$PlayerName),horiz=TRUE,cex.names=0.8,axis.lty=1,col=ifelse(PlayersDF$Position=='SG',"red",ifelse(PlayersDF$Position=="C","yellow",ifelse(PlayersDF$Position=="PG","blue",ifelse(PlayersDF$Position=="SF","orange",ifelse(PlayersDF$Position=="PF","green","gray"))))))

#Add a legend to the plot to show the colors that denote player position
legend("bottomright",legend=c("Shooting Guard","Center","Point Guard","Small Forward","Power Forward"),col=c("red","yellow","blue","orange","green"), pch=20)
